Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-17 snTomas.tex
finished the update to z=1.3457   * also ran spell-check

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
exinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from 14 lens models, representing a broad range of
methodologies and incorporating a variety of strong- and weak-lensing
constraints.  The models are collectively fairly accurate: 6 of the
models deliver median magnifications that are consistent with the
measured $\mu$ to within the 1$\sigma$ uncertainties.  However, the
models are systematically biased to higher values: \change{all but
one} of the tested models yield a median magnification higher than the
observed $\mu$, and \change{6 of the models are $\geq1.5\sigma$}
above the measured magnification.  We evaluate possible causes for
this mild bias, considering both modeling errors and
misinterpretation of the SN data. Although we find no single
explanation that can resolve this, we \change{ highlight several ways
that lens models can deliver more realistic uncertainties, such as
incorporating errors from assumed cosmological parameters and allowing
greater flexibility in the form of the cluster mass distribution.  A
more robust evaluation of systematic biases could be achieved with
further cooperative lens model comparisons using a larger sample of
lensed Type Ia SNe and cluster lensing simulations.}
\end{abstract}

@@

SN \tomas\ was discovered in Hubble Space Telescop (\HST) observations
with the Advanced Camera for Surveys (ACS) in the F606W and F814W
bands (V and i), collected on UT 2014 May 15 as part of the Hubble
Frontier Fields (HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying parallel fields.  Each
field is observed in 3 optical bands (ACS F435W, F606W and F814W) and
4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}, and the
locations of the cluster center, the SN, and the presumed host galaxy
are given in Table~\ref{tab:Positions}.  The SN detection was made in
difference images constructed using template imaging of Abell 2744
from HST+ACS observations taken in 2009 (PI:Dupke, HST-PID:11689).

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.3$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K features.
This \ion{Ca}{2} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in Section~\ref{sec:PhotometricClassification},
this spectral classification of SN \tomas\ is reinforced by the
photometric information, which also supports classification as a Type
Ia SN at $z\approx1.3$.

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  From this analysis we find
results that are consistent with the SNID fits \change{and the host
galaxy redshift}: $z=1.31\pm0.03$ and a phase of $0\pm3$ rest-frame
days. The best-fitting spectral template is the normal Type Ia SN
2012cg, shown in the top panel of Figure~\ref{fig:SpecFit}, which has
$\chi^2$ per degree of freedom $\nu$ equal to 42.1/87.  \change{If we
fix the redshift at $z=1.3457$ then we find the best fit is from the
normal Type Ia SN 2014J, with $\chi^2/\nu=51.8/87$.}

Next, we repeat the fitting, but without any warping of the templates
to account for differences in the continuum shape.  In this iteration
we only allow each template SED to be scaled in flux coherently at all
wavelengths.  As shown in Figure~\ref{fig:SpecFit}, we again find that when the \change{redshift is allowed as a free parameter} 
the \tomas\ SED is best matched by a normal Type Ia SN (SN 2011fe) at
redshift $z=1.31\pm0.02$.  \change{Fixing the redshift to 
$z=1.3457$ of the host, we also find the best match is from SN 2011fe, with
$\chi^2/\nu=133.7/91$. }

Without the continuum warping, an alternative
fit also arises: the fast-declining (91bg-like) Type Ia SN 2011iv at
$z=0.98\pm0.01$. Formally, this match provides a slightly better fit
to the unwarped \tomas\ spectrum, although the fit is notably poor at
$\sim8700$\AA\ where the most significant absorption feature is
found. Furthermore, a redshift $z\sim1$ is at odds with the host
galaxy photo-z ($1.5\pm0.2$), and we will see in the following section
that the photometric data is also incompatible with a Type Ia SN at
$z\sim1.0$.

@@

The marginal posterior distribution in redshift for the Type Ia model
is sharply peaked at $z=1.35\pm0.02$, which is fully consistent with
the photo-$z$ of the presumed host galaxy ($z=1.5\pm0.2$), and
$<2\sigma$ from the spectroscopic redshift of $z=1.31\pm0.01$ derived
in Section~\ref{sec:SpecRedshift}.  The time of peak brightness is
also tightly constrained at $t_{\rm pk}=56816.3\pm0.3$, which means
the spectroscopic observations were collected within 2 rest-frame days
of the epoch of peak brightness, consistent with our spectroscopic
analysis.

@@

With the spectroscopic type and redshift securely defined as a normal
Type Ia SN at $z=1.31$, we now turn to fitting the light curve with
Type Ia templates to measure the distance modulus.  Here we use two
independent light curve fitters: the SALT2 model described
above and the MLCS2k2 model \citep{Jha:2007}.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/snTomas_lightcurve_fit_fluxAB25}
\caption{ Type Ia light curve fits to SN \tomas\ using the MLCS2k2
(top row) and SALT2 (bottom row) fitters. The model redshifts are set to
$z=1.33\pm0.02$, encompassing both redshift values as determined from
spectroscopic and photometric constraints.  Solid lines denote the
best-fit model and shaded lines show the range allowed by 1-$\sigma$
uncertainties on the model parameters. 
Observed fluxes are shown as diamonds, scaled to an AB magnitude
zero point of 25. Error bars are plotted, but most are commensurate
with the size of the points. The left-most panel includes observations
in the F435W and F606W filters, although these were not used for the
fit, as they are bluer than the minimum wavelength for the model. The
lower axis marks time in observer frame days, while the top axis shows
the time in the rest frame relative to the epoch of peak brightness.
\label{fig:LightCurveFits} }
\end{center}
\end{figure*}

With both fitters we find light curve shape and color parameters that
are fully consistent with a normal Type Ia SN, regardless of whether
we adopt the spectroscopic redshift $z=1.31$ from
Section~\ref{sec:SpecRedshift} or the photometric redshift $z=1.35$
from Section~\ref{sec:PhotometricClassification}.  For SALT2, with the
redshift range set to $1.33\pm0.02$, we find a light curve shape
parameter of $x_1=0.164\pm0.199$ and a color parameter of
$c=-0.115\pm0.025$, yielding a $\chi^2$ value of 48.4 for 36 degrees
of freedom, $\nu$.  With the MLCS2k2 fitter the best-fit shape
parameter is $\Delta=-0.033\pm0.083$ and the color term is
$A_{\rm V}=0.014\pm0.037$, giving $\chi^2/\nu=23.9/36$.

@@

As in \P14, we derive a distance modulus\footnote{We use 'dm' to
indicate the distance modulus to avoid confusion, reserving the symbol
$\mu$ to refer to the lensing magnification. This 'dm' is a standard
distance modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$
is the luminosity distance in Mpc.} while ignoring the
effects of gravitational lensing, so that we can later use it to
derive a direct measurment of the magnification.  From the SALT2 fit
we get

@@

\noindent  Here the parameters for light curve shape $s$ and color $C$ 
correspond to the SiFTO light curve fitter \citep{Conley:2008}, so we
first use the formulae from \citet{Guy:2010} to convert from SALT2
($x_1$ and $c$) into the equivalent SiFTO parameters.  We also add an
offset of 0.27 mag to the value of $m_B^*$ returned by SNANA, in order
to match the arbitrary normalization of the SALT2 fitter used
by \citet{Guy:2010} and \citet{Sullivan:2011}.  This conversion allows
us to adopt values for the constants $M$, $\alpha$, and $\beta$
from \citet{Sullivan:2011}, which have been calibrated using 472 SNe
from the SNLS3 sample \citep{Conley:2011}: $M=-19.12\pm0.03$,
$\alpha=1.367\pm0.086$, and $\beta=3.179\pm0.101$.

@@

The SNANA version of the MLCS2k2 fitter returns a value for the
distance modulus (dm$_{\rm MLCS2k2}$) that has an arbitrary zero point
offset relative to the SALT2 distances (dm$_{\rm SALT2}$).  To put the
two distances onto the same reference frame we add a zeropoint
correction of 0.20 mag to the MLCS2k2 distances as
in \P14.  This correction was derived by applying both
fitters to a sample of Type Ia SNe from the SDSS
survey \citep{Holtzman:2008,Kessler:2009b}, with the extinction law
$R_V$ fixed at 1.9.

@@

Final values for the SN distance modulus are shown in
Table~\ref{tab:MeasuredMagnification}, adopting the spectroscopic
redshift ($z=1.31$).  In addition to modifying the distance modulus
uncertainty for SN \tomas, we also add $\sigma_{\rm int}=0.08$ mag in
quadrature to the uncertainty for every SN in the control sample. Note
however that the uncertainty on the control sample value at $z=1.31$
is only 0.05 mag, smaller than the intrinsic dispersion because it
reflects only our measurement error on the mean distance modulus of
the population.  The distance moduli derived from the SALT2 and
MLCS2k2 light curve fitters are fully consistent within the
uncertainties.

@@

It is now an accepted practice in cosmological analyses using Type Ia
SN to apply a correction to the luminosity of each SN based on the
stellar mass of its host galaxy. Typically this is described as a
simple bifurcation of the SN population: SNe that appear in more
massive hosts are observed to be $\sim$0.08 mag brighter (after
corrections for light curve shape and color) than SNe in low-mass
hosts \citep{Kelly:2010,Sullivan:2010}.  The dividing line for this
purely empirical ``mass step'' correction is generally set around
$10^{10}$\Msun, although this is somewhat arbitrary \citep[see
e.g.][]{Betoule:2014}.  This happens to be very close to the
SN \tomas\ host galaxy mass, which we have measured to be
$10^{10.135}$ \Msun\ based on Eq. 7 of \citet{Taylor:2011}, using the
{\it BPZ} SED fit to define rest-frame optical magnitudes.

The physical mechanism that drives the mass step is not yet
understood, but may be related to the metallicity or age of the SN
progenitor systems. In either case, it is reasonable to expect that
the significance of this effect decreases with redshift, as metal-rich
passive galaxies become much less common at $z>1$.  Indeed, when the
size of the mass step correction is allowed to vary with redshift,
there is no significant evidence that a non-zero mass step is required
at $z>1$ \citep{Rigault:2013,Shafer:2014,Betoule:2014}.  Given the
absence of a clear physical model and the lack of empirical support
for a high-$z$ mass step, we do not apply any correction to SN \tomas\ 
or the control sample.

@@

To measure the lensing magnification, we would like to avoid
introducing systematic uncertainties inherent to any assumed
cosmological model \citep[e.g.][]{Nordin:2014}.  To that end, we
follow \P14\ and define the magnification by comparing
the measured distance modulus of SN \tomas\ against an average
distance modulus derived from a ``control sample'' of unlensed Type Ia
SN at similar redshift.  This allows us to make only the minimal
assumption that the redshift-distance relationship for Type Ia SN is
smooth and approximately linear over a small redshift span, which
should be true for any plausible cosmological model.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. Solid vertical lines show the constraints from SN \tomas\
derived in Section~\ref{sec:Magnification} using the SALT2 (red) and
MLCS2k2 (blue) fitters, with shaded regions marking the total
uncertainty for each. Markers with horizontal error bars show the
median magnification and 68\%\ confidence region from each of the 14
lensing models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  Dashed lines mark the unweighted mean of each model
family and shaded regions show the standard error of the mean.
\label{fig:lensingtest} }
\end{center}
\end{figure}

Table~\ref{tab:PredictedMagnifications} lists the \change{11} models,
giving for each one the predicted magnification and uncertainty for a
source at \change{$z=1.3457$} and at the position of \tomas.  These models
represent a broad sampling of the techniques and assumptions that can
be applied to the modeling of mass distributions in galaxy clusters. A
rigorous comparison of these disparate lens modeling techniques is
beyond the scope of this work. Table~\ref{tab:PredictedMagnifications}
includes a brief description of each model, but for a complete
discussion of the methodology the reader is referred to the listed
references.

In Figure~\ref{fig:lensingtest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:Magnification}.  This comparison shows that these 14
models are largely consistent with each other. Naively treating each
model as an independent prediction for the magnification (and ignoring
their quoted uncertainties), one would get a mean for the full sample
of $\mu=2.51\pm0.34$.  This is separated from the observed SN
magnification by $\delta\mu/\mu=27\%$, which is only a 1.2$\sigma$
difference.

@@

Figure~\ref{fig:lensingtest} also separates the models based on the
scope of input data constraints used.  Ten of the models rely only
on strong-lensing constraints (plotted as circles), while the other
four also use weak-lensing measurements (diamonds).  Ten of the
models were constructed using only pre-HFF data (open symbols), and
four others added new multiply-imaged systems and redshifts.

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  Six of the tested models
(CATS, GLAFIC, Zitrin-NFW, Bradac(v2), Merten, Diego) report a median
value that falls within the 1$\sigma$ uncertainty range of the
measured magnification, and two of the free-form models (Williams,
Diego) report uncertainties that overlap the central $\mu$ as measured
from both of the SN light curve fitters.  Furthermore, the scatter
amongst models provides a reasonable estimate of the magnification
uncertainty.

@@

However, beyond this first-order agreement, there is a systematic bias
apparent. All of the lens models return median magnifications
that are {\it higher} than the observed value, and half of
the models are discrepant by more than $1\sigma$.  It is important to
emphasize that SN \tomas\ only samples a single sight-line through the
cluster, so any conclusions to be drawn from this analysis are
necesssarily limited.  Nevertheless, a systematic shift common to all
models is surprising, given the wide range of modeling strategies,
input data, and physical assumptions represented by this set of
models.  In the following subsections we examine possible explanations
for this small but apparently universal bias.

@@

We tested this hypothesis using the CATS-SL(v2) model by allowing the mass
of the nearby cluster member galaxy to vary as a free parameter in the
model. We found that the change in the SN \tomas\ magnification
prediction was less than $\Delta\mu=0.1$.  This additional dispersion
is already included in the uncertainties quoted for that model in
Table~\ref{tab:PredictedMagnifications} and
Figure~\ref{fig:lensingtest}.  Furthermore, an erroneous M/L
value for the nearby cluster member galaxy could not explain the
systematic shift of all lens models, because some do not incorporate
cluster member galaxies into their constraints at all (the free-form
Bradac, Williams and Merten models).

@@

With two versions of the CATS model, we also have a more direct test
of the effect of introducing weak lensing constraints.  The initial
CATS model uses only strong-lensing constraints, and gives
$\mu=2.27^{+0.05}_{-0.04}$, slightly higher than the measured value.
A second iteration of this model, labeled here as CATS-SL+WL, used the
same strong-lensing features, but added in weak lensing constraints.
The revised magnification of $\mu=2.62\pm0.18$ is further from the
measured value, which serves to reinforce the conclusion that adding
weak lensing constraints cannot, by itself, resolve this bias.

>>>>

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from 14 lens models, representing a broad range of
methodologies and incorporating a variety of strong- and weak-lensing
constraints.  The models are collectively fairly accurate: 6 of the
models deliver median magnifications that are consistent with the
measured $\mu$ to within the 1$\sigma$ uncertainties.  However, the
models are systematically biased to higher values: \change{all but
one} of the tested models yield a median magnification higher than the
observed $\mu$, and \change{6 of the models are $\geq1.5\sigma$}
above the measured magnification.  We evaluate possible causes for
this mild bias, considering both modeling errors and
misinterpretation of the SN data. Although we find no single
explanation that can resolve this, we \change{ highlight several ways
that lens models can deliver more realistic uncertainties, such as
incorporating errors from assumed cosmological parameters and allowing
greater flexibility in the form of the cluster mass distribution.  A
more robust evaluation of systematic biases could be achieved with
further cooperative lens model comparisons using a larger sample of
lensed Type Ia SNe and cluster lensing simulations.}
\end{abstract}

@@

SN \tomas\ was discovered in Hubble Space Telescope (\HST) observations
with the Advanced Camera for Surveys (ACS) in the F606W and F814W
bands (V and i), collected on UT 2014 May 15 as part of the Hubble
Frontier Fields (HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying parallel fields.  Each
field is observed in 3 optical bands (ACS F435W, F606W and F814W) and
4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}, and the
locations of the cluster center, the SN, and the presumed host galaxy
are given in Table~\ref{tab:Positions}.  The SN detection was made in
difference images constructed using template imaging of Abell 2744
from HST+ACS observations taken in 2009 (PI:Dupke, HST-PID:11689).

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K features.
This \ion{Ca}{2} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in Section~\ref{sec:PhotometricClassification},
this spectral classification of SN \tomas\ is reinforced by the
photometric information, which also supports classification as a Type
Ia SN at $z\approx1.35$.

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  \change{With the redshift fixed
at $z=1.3457$ we find the best fit is from the normal Type Ia SN
2014J, with a $\chi^2$ per degree of freedom $\nu$ of
$\chi^2/\nu=51.8/87$ shown in the top panel of
Figure~\ref{fig:SpecFit}.  Allowing the redshift as an additional free
parameter, we still find results that are consistent with the SNID
fits and $<2\sigma$ from the host galaxy redshift: $z=1.31\pm0.02$ and
a phase of $0\pm3$ rest-frame days. The best-fitting spectral template
in this case is the normal Type Ia SN 2012cg at $z=1.31$ with
$\chi^2/\nu=42.1/87$.}

Next, we repeat the fitting, but without any warping of the templates
to account for differences in the continuum shape.  In this iteration
we only allow each template SED to be scaled in flux coherently at all
wavelengths.  As shown in Figure~\ref{fig:SpecFit}, we again find that
when the \change{redshift is allowed as a free parameter the \tomas\
SED is best matched by a normal Type Ia SN (SN 2011fe) at redshift
$z=1.31$ with $\chi^2/\nu=102.9/90$.  Fixing the redshift to the
redshift $z=1.3457$ of the host, we also find the best match is from
SN 2011fe, with $\chi^2/\nu=133.7/91$. }

Without the continuum warping, an alternative
fit also arises: the fast-declining (91bg-like) Type Ia SN 2011iv at
$z=0.98\pm0.01$. Formally, this match provides a slightly better fit
to the unwarped \tomas\ spectrum, although the fit is notably poor at
$\sim8700$\AA\ where the most significant absorption feature is
found. Furthermore, a redshift $z\sim1$ is at odds with the host
galaxy redshift ($z=1.3457$), and we will see in the following section
that the photometric data is also incompatible with a Type Ia SN at
$z\sim1.0$.

@@

The marginal posterior distribution in redshift for the Type Ia model
is sharply peaked at $z=1.35\pm0.02$, which \change{is fully
consistent with the redshift of the presumed host galaxy ($z=1.3457$)
as well as the redshift of $z=1.31\pm0.02$ derived from the SN
spectrum in Section~\ref{sec:SpecRedshift}.  The time of peak
brightness is also tightly constrained at $t_{\rm pk}=56816.3\pm0.3$,
which means the HST grism observations were collected within 2
rest-frame days of the epoch of peak brightness -- also consistent
with our spectroscopic analysis.}

@@

With the type and redshift securely defined as a normal
Type Ia SN at $z=1.3457$, we now turn to fitting the light curve with
Type Ia templates to measure the distance modulus.  Here we use two
independent light curve fitters: the SALT2 model described
above and the MLCS2k2 model \citep{Jha:2007}.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/snTomas_lightcurve_fit_fluxAB25}
\caption{ Type Ia light curve fits to SN \tomas\ using the MLCS2k2
(top row) and SALT2 (bottom row) fitters. The model redshifts are set to
$z=1.3457$ as determined from combined
spectroscopic and photometric constraints.  Solid lines denote the
best-fit model and shaded lines show the range allowed by 1-$\sigma$
uncertainties on the model parameters. 
Observed fluxes are shown as diamonds, scaled to an AB magnitude
zero point of 25. Error bars are plotted, but most are commensurate
with the size of the points. The left-most panel includes observations
in the F435W and F606W filters, although these were not used for the
fit, as they are bluer than the minimum wavelength for the model. The
lower axis marks time in observer frame days, while the top axis shows
the time in the rest frame relative to the epoch of peak brightness.
\label{fig:LightCurveFits} }
\end{center}
\end{figure*}

With both fitters we find light curve shape and color parameters that
are fully consistent with a normal Type Ia SN.  For SALT2, with the
\change{redshift fixed at $z=1.3457$, we find a light curve shape
parameter of $x_1=0.135\pm0.199$ and a color parameter of
$c=-0.127\pm0.025$, yielding a $\chi^2$ value of 45.0 for 36 degrees
of freedom, $\nu$.  With the MLCS2k2 fitter the best-fit shape
parameter is $\Delta=-0.082\pm0.070$ and the color term is
$A_{\rm V}=0.011\pm0.025$, giving $\chi^2/\nu=22.8/36$.}

@@

As in \citealt{Patel:2014}, we derive a distance modulus (dm)\footnote{We use 'dm' to
indicate the distance modulus to avoid confusion, reserving the symbol
$\mu$ to refer to the lensing magnification. This 'dm' is a standard
distance modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$
is the luminosity distance in Mpc.} while ignoring the
effects of gravitational lensing, so that we can later use it to
derive a direct measurement of the magnification.  \change{The MLCS2k2 fitter returns $\rm{dm}_{\rm MLCS2k2}$ directly, as it is defined as one of the free parameters in the model. To derive dm from} the SALT2 fit
we use

@@

\noindent  Here the parameters for light curve shape $s$ and color $C$ 
correspond to the SiFTO light curve fitter \citep{Conley:2008}, so we
first use the formulae from \citet{Guy:2010} to convert from SALT2
($x_1$ and $c$) into the equivalent SiFTO parameters.  We also add an
offset of 0.27 mag to the value of $m_B^*$ returned by SNANA, in order
to match the arbitrary normalization of the SALT2 fitter used
by \citet{Guy:2010} and \citet{Sullivan:2011}.  \change{This conversion from SALT2 to SiFTO is necessary, as it} allows
us to adopt values for the constants $M$, $\alpha$, and $\beta$
from \citet{Sullivan:2011}, which have been calibrated using 472 SNe
from the SNLS3 sample \citep{Conley:2011}: $M=-19.12\pm0.03$,
$\alpha=1.367\pm0.086$, and $\beta=3.179\pm0.101$.

@@

The SNANA version of the MLCS2k2 fitter returns a value for the
distance modulus (dm$_{\rm MLCS2k2}$) that has an arbitrary zero point
offset relative to the SALT2 distances (dm$_{\rm SALT2}$).  To put the
two distances onto the same reference frame we add a zeropoint
correction of 0.20 mag to the MLCS2k2 distances as
in \citealt{Patel:2014}.  This correction was derived by applying both
fitters to a sample of Type Ia SNe from the SDSS
survey \citep{Holtzman:2008,Kessler:2009b}, with the extinction law
$R_V$ fixed at 1.9.

@@

Final values for the SN distance modulus are shown in
Table~\ref{tab:MeasuredMagnification}, adopting the spectroscopic
redshift ($z=1.3457$).  In addition to modifying the distance modulus
uncertainty for SN \tomas, we also add $\sigma_{\rm int}=0.08$ mag in
quadrature to the uncertainty for every SN in the control sample. Note
however that the uncertainty on the control sample value at $z=1.3457$
is only 0.06 mag, smaller than the intrinsic dispersion because it
reflects only our measurement error on the mean distance modulus of
the population.  The distance moduli derived from the SALT2 and
MLCS2k2 light curve fitters are fully consistent within the
uncertainties.

@@

It is now an accepted practice in cosmological analyses using Type Ia
SN to apply a correction to the luminosity of each SN based on the
stellar mass of its host galaxy. Typically this is described as a
simple bifurcation of the SN population: SNe that appear in more
massive hosts are observed to be $\sim$0.08 mag brighter (after
corrections for light curve shape and color) than SNe in low-mass
hosts \citep{Kelly:2010,Sullivan:2010}.  The dividing line for this
purely empirical ``mass step'' correction is generally set around
$10^{10}$\Msun, although this is somewhat arbitrary \citep[see
e.g.][]{Betoule:2014}.  This happens to be very close to the
SN \tomas\ host galaxy mass, which we have measured to be
$10^{10.135}$ \Msun\ based on Eq. 7 of \citet{Taylor:2011}, using the
{\it BPZ} SED fit (Section~\ref{sec:DiscoveryAndFollowup}) to define rest-frame optical magnitudes.

The physical mechanism that drives the mass step is not yet
understood, but may be related to the metallicity or age of the SN
progenitor systems.  In either case, the significance of this effect
should decrease with redshift, as metal-rich passive galaxies become
much less common at $z>1$ \change{\citep[see
e.g.][]{Rigault:2013,Childress:2014b}.}  Indeed, when the size of the
mass step correction is allowed to vary with redshift, there is no
significant evidence that a non-zero mass step is required at
$z>1$ \citep{Rigault:2013,Shafer:2014,Betoule:2014}.  Given the
absence of a clear physical model and the lack of empirical support
for a high-$z$ mass step, we do not apply any correction to SN \tomas\
or the control sample.

@@

To measure the lensing magnification, we would like to avoid
introducing systematic uncertainties inherent to any assumed
cosmological model \citep[e.g.][]{Nordin:2014}.  To that end, we
follow \citealt{Patel:2014} and define the magnification by comparing
the measured distance modulus of SN \tomas\ against an average
distance modulus derived from a ``control sample'' of unlensed Type Ia
SN at similar redshift.  This allows us to make only the minimal
assumption that the redshift-distance relationship for Type Ia SN is
smooth and approximately linear over a small redshift span, which
should be true for any plausible cosmological model.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. Solid vertical lines show the constraints from SN \tomas\
derived in Section~\ref{sec:Magnification} using the SALT2 (red) and
MLCS2k2 (blue) fitters, with shaded regions marking the total
uncertainty for each. Markers with horizontal error bars show the
median magnification and 68\%\ confidence region from each of the 14
lensing models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  The grey dashed line marks the unweighted mean for the
set of all models, at $\mu=2.53$.
\label{fig:lensingtest} }
\end{center}
\end{figure}

Table~\ref{tab:PredictedMagnifications} lists the 14 models,
giving for each one the predicted magnification and uncertainty for a
source at \change{$z=1.3457$} and at the position of \tomas.  These models
represent a broad sampling of the techniques and assumptions that can
be applied to the modeling of mass distributions in galaxy clusters. A
rigorous comparison of these disparate lens modeling techniques is
beyond the scope of this work. Table~\ref{tab:PredictedMagnifications}
includes a brief description of each model, but for a complete
discussion of the methodology the reader is referred to the listed
references.

In Figure~\ref{fig:lensingtest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:Magnification}.  This comparison shows that these 14
models are largely consistent with each other. Naively treating each
model as an independent prediction for the magnification (and ignoring
their quoted uncertainties), one would get a mean for the full sample
of \change{$\mu=2.53\pm0.38$.  This is separated from the observed SN
magnification by $\delta\mu/\mu=25\%$, which is just over a
1$\sigma$ difference.}

@@

Figure~\ref{fig:lensingtest} also separates the models based on the
scope of input data constraints used.  Ten of the models rely only on
strong-lensing constraints (plotted as circles), while the other four
also use weak-lensing measurements (diamonds).  Nine of the models
were constructed using only pre-HFF data (open symbols), while five
others added new multiply-imaged systems and redshifts.

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  Six of the tested models
(CATSv1, GLAFIC, Zitrin-NFW, Bradac(v2), Merten, Diego) report a median
value that falls within the 1$\sigma$ uncertainty range of the
measured magnification, and two of the free-form models (Williams,
Diego) report uncertainties that overlap the central $\mu$ as measured
from both of the SN light curve fitters.  Furthermore, the scatter
amongst models provides a reasonable estimate of the magnification
uncertainty.

@@

However, beyond this first-order agreement, there is a small
systematic bias apparent. All of the lens models \change{except one}
return median magnifications that are {\it higher} than the observed
value, and \change{six of the models are discrepant by more than $1.5\sigma$.}
It is important to emphasize that SN \tomas\ only samples a single
sight-line through the cluster, so any conclusions to be drawn from
this analysis are necessarily limited.  Nevertheless, a systematic
shift common to all models is surprising, given the wide range of
modeling strategies, input data, and physical assumptions represented
by this set of models.  In the following subsections we examine
possible explanations for this small but \change{nearly} universal
bias.

@@

We tested this hypothesis using the CATS(v2) model by allowing the mass
of the nearby cluster member galaxy to vary as a free parameter in the
model. We found that the change in the SN \tomas\ magnification
prediction was less than $\Delta\mu=0.1$.  This additional dispersion
is already included in the uncertainties quoted for that model in
Table~\ref{tab:PredictedMagnifications} and
Figure~\ref{fig:lensingtest}.  Furthermore, an erroneous M/L
value for the nearby cluster member galaxy could not explain the
systematic shift of all lens models, because some do not incorporate
cluster member galaxies into their constraints at all (the free-form
Bradac, Williams and Merten models).

@@

With two versions of the CATS model, we also have a more direct test
of the effect of introducing weak lensing constraints.  The initial
CATS(v1) model uses only strong-lensing constraints, and gives
$\mu=2.29^{+0.05}_{-0.04}$, slightly higher than the measured value.
A second iteration of this model, labeled here as CATS(v1sw), used the
same strong-lensing features, but added in weak lensing constraints.
The revised magnification of $\mu=2.62\pm0.18$ is further from the
measured value, which serves to reinforce the conclusion that adding
weak lensing constraints cannot, by itself, resolve this bias.
